Inferring Continuous and Discrete Population Genetic Structure Across Space

### 2017-08-15 writeup/conStruct.tex

Several approaches have been used as model choice criteria 
for the number of clusters in population genetic data, including: 
the use of information theoretic criteria \citep[e.g.][]{ADMIXTURE}, such as AIC or BIC;
comparisons of the likelihood of the data across different values of $K$, 
with various criteria on how to choose a single value (e.g., \citep{Evanno2005});
or comparisons of the marginal likelihood, 
generated either via various approximations \citep[e.g.][]{STRUCTURE}
or via a more rigorous and computationally intensive approach such as thermodynamic integration \citep{verity_nichols2016}
or inference using a Dirichlet process prior \citep{huelsenbeck2007inference}
See \cite{verity_nichols2016} for a discussion of these approaches and comparison
between several methods.
\plr{I wonder if we can just cite \cite{verity_nichols2016} rather than this laundry list here?}

>>>>

Several approaches have been used as model choice criteria 
for the number of clusters in population genetic data, including: 
the use of information theoretic criteria \citep[e.g.][]{ADMIXTURE}, such as AIC or BIC;
comparisons of the likelihood of the data across different values of $K$, 
with various criteria on how to choose a single value (e.g., \citep{Evanno2005});
or comparisons of the marginal likelihood, 
generated either via various approximations \citep[e.g.][]{STRUCTURE}
or via a more rigorous and computationally intensive approach such as thermodynamic integration \citep{verity_nichols2016}
or inference using a Dirichlet process prior \citep{huelsenbeck2007inference}
See \cite{verity_nichols2016} for a discussion of these approaches and comparison
between several methods.

\gc{IS PREDICTION accuracy the right phrase?}
We choose to take a cross-validation approach \citep[similar in spirit to][]{ADMIXTURE_xval}.
We estimate the posterior on the parameters of the model using a ``training" partition of the data 
(in practice the training set is a random subset of $90\%$ of the loci). 
We then calculate the likelihood of the ``testing" subset (the remaining $10\%$ of the loci), 
averaged over the posterior of the parameters. 
We repeat this procedure a number of times, 
take an average of the relative log-likelihoods over the runs at each $K$, 
and select the model with the significantly best predictive accuracy
(see Section \ref{Xvalidation} for more on our cross-validation procedure).
This approach rewards models with larger values of $K$, 
which increase the likelihood of the training partition, 
as they fit the data better,
but the improvement in the likelihood will plateau (or peak), 
as above a certain $K$ the model is fitting noise specific to the training data 
and therefore not fitting the testing set any better.
\plr{I wonder if we can just cite \cite{verity_nichols2016} rather than this laundry list here?}
